---
title: "Webscraping IMBD Ratings of The Office Episodes"
output: html_notebook
---

The Office is an extremely popular American Sitcom. The show was based on a 
British show of the same name, and features a "mockumentary" style in which a 
crew documents the lives of ordinary American workers at a paper company. The 
show centers on the Regional Manager "Michael Scott" (Steve Carell). Michael is 
a clueless boss and is convinced that all his employees love him. The employees, 
however, are not amused by Michael's absurd and sometimes offensive antics. 

Steve Carell left the Office in Season 7. The show aired for 2 more seasons after
his departure. Using data that I webscraped from IMBD I am interested in how Steve
Carrells departure from the show effected the ratings. 

```{r}
rm(list = ls())

library(rvest)
library(ggplot2)

```

Using the rvest library, I scrape data from IMBD, extracting the episode name, year, and rating.

```{r}
# Scraping Imbd ratings for "The Office" from the Web
url <- "https://www.imdb.com/search/title/?series=tt0386676&view=simple&count=250&sort=user_rating,desc"
webpage <- read_html(url)

# Using Rvest and CSS selector extension I locate the node in the html and extract the text

# Get the rating
ratings <- html_elements(webpage, ".col-imdb-rating")
as.numeric(html_text(ratings, trim = TRUE)) #convert it to numeric data

# Getting the episode name
episode <- html_elements(webpage, ".unbold+ a")
html_text(episode)

# Getting the year 
year <- html_elements(webpage, ".unbold~ .text-muted")
html_text(year)


```
I then put the scraped information in a dataframe suitable for visualization and analysis. 

```{r}
# Put the scraped elements in a new data frame
office_data <- data.frame(html_text(episode), 
                            as.numeric(html_text(ratings, trim = TRUE)),
                            html_text(year))

# give sensible names to the dat frame
names(office_data) <- c("episode", "rating", "year")
```

IMBD site does not have info on whether Michael Scott was in a particular episode.
Using my own knowledge of the show, I created a variable that tracks whether Michael
was in the show. Michael was in every episode for the first 6 seasons. These seasons
correspond to the years 2005 - 2010. Michael was in most of the episodes for season 7,
and returned for the finale in season 9. Because these episodes do not cleanly 
correspond to years I created a short list of Michael Episodes that were not 
captured by the years.

```{r}
# Create a variable tracking if the character "Michael Scott" was in the episode

office_data$michael_scott <- "No"

# Create a character vector for the years and episodes the Micheal was in
michael_scott_years <- c("(2005)", "(2006)", "(2007)", "(2008)", "(2009)", "(2010)")
michael_scott_episodes <- c("Finale", "Goodbye, Michael", "Garage Sale", "Threat Level Midnight",
                            "Michael's Last Dundies", "The Search", "PDA", "Ultimatum",
                            "Training Day", "The Seminar", "Todd Packer")


# I index the data checking to see if the row was in one of these character vectors
# and change the Michael Scott status to "Yes" accordingly 
office_data$michael_scott[office_data$year %in% michael_scott_years] <- "Yes"
office_data$michael_scott[office_data$episode %in% michael_scott_episodes] <- "Yes"
```

First I am interested in the overall rating based on Michael Scott's character.
Statistical analysis reveals that episodes with Michael in them had an average
rating of 8.26 and that episodes without had an average of 7.55. This is nearly 
a 1 point difference on a scale of 10, and is statistically significant p < 0.001.

```{r}
ggplot(data = office_data, mapping = aes(x = michael_scott, y = rating)) +
  geom_boxplot() +
  xlab("Michael Scott in the Episode?") +
  ylab("IMBD Rating") +
  ggtitle("The Office Episode Ratings")

t.test(office_data$rating ~ office_data$michael_scott)


```

I was also interested in how the show fared overtime. Many fans - myself included - 
consider seasons 2-4 to be the golden era of The office. Season 8, the first season
without Michael, is often considered to be a lowpoint in the series. 


```{r}
ggplot(data = office_data, mapping = aes(x = year, y = rating)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013"), name = "Year") +
  ylab("IMBD Ratng") +
  ggtitle("Ratings of The Office by Year")
```
Although the years do not perfectly correspond to the seasons (Most season span over 
two calendar years) this graph reflects the traditional wisdom. Years 2006 - 2009, which 
encompass seasons 2-5 experience the highest average ratings. Meanwhile year 2012, which
had most of season 8 episodes, is a low point on the graph. 

While The Office endend on a strong note - the finale was one of the highest rated episodes -
the show did suffer from the departure of it's main character. Using web-scraping methods
we can see that the episodes and years without Michael Scott had significantly lower
ratings on average. 



